---
title: "Final Project"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(readxl)
library(leaps)
library(ISLR)                          
library(psych)                         # R data description
library(boot)                          # For cv.glm()
library(dplyr)
library(car)
library(MPV)
library(HH)
```

```{r}
hospital = read_excel("./data/GHProject_Dataset.xlsx") %>%
  clean_names()
```

```{r}
data = hospital[!duplicated(hospital$patientid), ] 
#delete duplicated patientid, patientid and visitid is now one-to-one

data = 
  data %>%
  separate(admitdtm, into = c("weekday", "day_month", "year"), sep = ",") %>%
  mutate(day_month = trimws(day_month), year = trimws(year)) %>%
  separate(day_month, into = c("month_1", "date"), sep = " ") %>%
  mutate(month = match(month_1, month.name)) %>%
  mutate(date_of_admit = paste(month, date, year, sep = "/"),
         date_of_admit = as.Date(date_of_admit, "%m/%d/%Y")) %>%
  arrange(date_of_admit) %>%
  dplyr::select(-weekday, -month_1, -month, -year, -date) %>%
  dplyr::select(patientid, visitid, loshours, losdays2, date_of_admit, everything())
# til now just change the date format into yy-mm-dd
# I haven't deleted any important variable columns just in case you guys need them while modeling and I have to tranform each variable to see its skewness.

#Delete the conflict that a patient stays more than 30 days and with asmission into the hoppital within past 30 days.
### NOT sure about this!!! because it's saying readmit..after doing so, it deleted 6 entries.
data = data[!(data$is30dayreadmit == 1 & data$losdays2 > 30),]

###########################descriptive statistics of each variable############################
########continuous variables (9)
data_subset1 = data %>%
  dplyr::select(losdays2, ageyear, bmi, 
         bpsystolic, o2sat, temperature, 
         heartrate, respirationrate, bpdiastolic)

table_pre = map(data_subset1,summary) # the descriptive data shows bmi column has a lot of NA's

#create a table that combines the summary of the 9 continuous variables
knitr::kable(bind_rows(
  c(variable = "Losdays2",table_pre[[1]]),#good to do log transform
  c(variable = "AgeYear", table_pre[[2]]),#quite normal
  c(variable = "bmi",table_pre[[3]]),#688 missing values and long tail, 3.1 min and 123 max, outlier/error
  c(variable = "bpsystolic", table_pre[[4]]),
  c(variable = "o2sat", table_pre[[5]]), # 10 observation below 90, 80 observation above 100, 2% of the data
  c(variable = "temperature", table_pre[[6]]), #3 less than 30 and 2 more than 50
  c(variable = "heartrate", table_pre[[7]]), # 2 more than 240 (rare heart rate or caused by disease)
  c(variable = "respirationrate", table_pre[[8]]),
  c(variable = "bpdiastolic", table_pre[[9]])
))
 




#########categorical variables
#The Modified Early Warning Score (MEWS) 
#was 0-1=normal, 2-3=increase caution, 4-5=further deterioration, >5 immediate action required
table(data$mews) # using a table to check the distribution of mews. Skewed!
#make them into only 3 categories
#mews_new: 1=normal; 2=increase caution; 3 = further deterioration or immediate action required
data$mews_new <- ifelse(data$mews == 0|data$mews == 1, 1, ifelse(data$mews == 2|data$mews == 3, 2, 3))
table(data$mews_new) # a bit better


#Cindex ranks patients based on severity of comorbidity: 0=normal, 1-2=mild, 3-4=moderate and >5 = severe
table(data$cindex) # same as above - skewed
#make them into three categories
data$cindex_new <- ifelse(data$cindex == 0,"normal", ifelse(data$cindex == 1|data$cindex == 2,"mild","moderate to severe"))
table(data$cindex_new) # after balacing , better now.

#evisit -skewed
table(data$evisit) 
data$evisit_new <- ifelse(data$evisit==0|data$evisit==1, 1, 2)
table(data$evisit_new)
#evisit_new:1 = 0or1 emergency department visit; 2=more than 1 emergency department visits.


#ICU_Flag: 1=if during hospitalization, the patient had a visit in the ICU; 0=otherwise, -skewed
table(data$icu_flag)  

#gender  - ok
table(data$gender)

#race:   - skewed
table(data$race)

data$race_new <- ifelse(data$race=="White","White", ifelse(data$race=="African Amer/Black","African Amer/Black","Other race"))
table(data$race_new)
#only have 3 categories, AAB, White, other race.

#religion: recode them
data$religion[data$religion == "Angelican"] <- "Christian"
data$religion[data$religion == "Non Denominational"] <- "Christian"
data$religion[data$religion == "Catholic"] <- "Christian"
data$religion[data$religion == "Hebrew"] <- "Jewish"
data$religion[data$religion == "Mormon"] <- "Other"

table(data$religion)


#MaritalStatus: - skewed
table(data$maritalstatus)
data$marital_new <- ifelse(data$maritalstatus=="Civial Union"|data$maritalstatus=="Married"|data$maritalstatus=="Civial Union","Not single","single")
table(data$marital_new)


#InsuranceType:  -skewed
table(data$insurancetype)
data$insurance_new <- ifelse(data$insurancetype=="Private", "Private", "Public")
table(data$insurance_new)



#visualize dataset for each variable
attach(data)
par(mfrow=c(3,3))
hist(losdays2) #highly skewed
hist(ageyear)
hist(bmi)
hist(bpsystolic)
hist(o2sat)
hist(temperature)
hist(heartrate)
hist(respirationrate)
hist(bpdiastolic)

#try transformations of data

hist(log(data$losdays2)) #good to do log transformation

```

Wendy:

```{r}
# Replace "funny" characters
# Looking at the histograms and googling about the vital signs, we can see that there are some "funny" characters which are obviously unusual, so we decide to replace those with NA.

# We replace BMI lower than 12 or higher than 50
data$bmi[data$bmi < 12|data$bmi > 50] <- NA


# We replace oxygen saturation higher than 200
data$o2sat[data$o2sat > 200] <- NA

# Human's heart will stop beating if the body's temperature is lower than 24 or higher than 47
data$temperature[data$temperature < 24|data$temperature > 47] <- NA

# We replace heartrate higher than 220
data$heartrate[data$heartrate > 220] <- NA

# We replace respiration rate higher than 50
data$respirationrate[data$respirationrate > 50] <- NA

# We replace BPD lower than 40 or higher than 140
data$bpdiastolic[data$bpdiastolic < 40|data$bpdiastolic > 140] <- NA

# After removing "funny" values, we look at the distributions of the predictors again.
attach(data)
par(mfrow=c(3,3))
hist(ageyear)
hist(bmi)
hist(bpsystolic)
hist(o2sat)
hist(temperature)
hist(heartrate)
hist(respirationrate)
hist(bpdiastolic)
# We think there is no need to transform any predictors.
```

```{r}
# variable selection
# First, we look at the meaning of different variables. Obviously, patient id, visit id and  admitdtm has nothing to do with patient's lenghth of stay, so we exclude them from the model.
# Next, we perform simple linear regression for each variable we select.
data_update <- data %>%
   mutate(trans_losdays2 = log(losdays2)) %>%
   filter(icu_flag == 0) %>%
   dplyr::select(trans_losdays2, is30dayreadmit, cindex_new, evisit_new, ageyear, gender, race_new, religion, marital_new, insurance_new, respirationrate, bpdiastolic, o2sat, bpsystolic, temperature, heartrate, bmi, mews_new) %>%
   na.omit()
 

lm1 <- lm(trans_losdays2 ~ is30dayreadmit, data = data_update)
summary(lm1)
# is30dayreadmit is significant

lm2 <- lm(trans_losdays2 ~ cindex_new, data = data_update)
summary(lm2)
# cindex is significant

lm3 <- lm(trans_losdays2 ~ evisit_new, data = data_update)
summary(lm3)
# evisit is significant

lm4 <- lm(trans_losdays2 ~ ageyear, data = data_update)
summary(lm4)
# ageyear is significant

lm5 <- lm(trans_losdays2 ~ gender, data = data_update)
summary(lm5)
# gender is not significant

lm6 <- lm(trans_losdays2 ~ race_new, data = data_update)
summary(lm6)
# race is not significant

lm7 <- lm(trans_losdays2 ~ religion, data = data_update)
summary(lm7)
# religion is not significant

lm8 <- lm(trans_losdays2 ~ marital_new, data = data_update)
summary(lm8)
# marital status is significant

lm9 <- lm(trans_losdays2 ~ insurance_new, data = data_update)
summary(lm9)
# insurance type is significant

lm10 <- lm(trans_losdays2 ~ respirationrate, data = data_update)
summary(lm10)
# respiration rate is significant

lm11 <- lm(trans_losdays2 ~ bpdiastolic, data = data_update)
summary(lm11)
# bpdiastolic is significant

lm12 <- lm(trans_losdays2 ~ o2sat, data = data_update)
summary(lm12)
# oxygen saturation is not really significant

lm13 <- lm(trans_losdays2 ~ bpsystolic, data = data_update)
summary(lm13)
# bpsystolic is significant

lm14 <- lm(trans_losdays2 ~ temperature, data = data_update)
summary(lm14)
# temperature is not really significant

lm15 <- lm(trans_losdays2 ~ heartrate, data = data_update)
summary(lm15)
# heartrate is significant

lm16 <- lm(trans_losdays2 ~ bmi, data = data_update)
summary(lm16)
# bmi is significant

lm17 <- lm(trans_losdays2 ~ mews_new, data = data_update)
summary(lm17)
# mews is significant

# Then we exculde the variables which are non-significant
data_update <- data_update %>%
  dplyr::select(-race_new, -religion,-gender)
```

```{r}
# Perform stepwise and best
lm_model <- lm(trans_losdays2 ~ ., data = data_update)
step(lm_model, direction = "both", trace = FALSE)

best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model),
                        nvmax=length(model$coefficient)-1,...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


# Select the 'best' 1 model of all subsets
round(best(lm_model, nbest = 1), 4)

# Based on the results above, we choose model : 
# trans_losdays2 ~ is30dayreadmit + cindex_new + evisit_new + ageyear + marital_new + insurance_new + respirationrate +  bpdiastolic + bpsystolic + temperature + heartrate
```

```{r}
lm_select <- lm(trans_losdays2 ~ is30dayreadmit + cindex_new + evisit_new + ageyear + marital_new + insurance_new + respirationrate +  bpdiastolic + bpsystolic + temperature + heartrate, data = data_update)
summary(lm_select)
```

##David
####Residuals vs fitted values plot
```{r}
plot(fitted(lm_select), resid(lm_select), xlab = "Predicted/Fitted value", ylab = "Residual")
title("(a) Residual Plot for Y (trans_losdays2) ")
abline(0, 0)
```

#### Quantile - Quantile plot (QQ-plot)
```{r}
qqnorm(resid(lm_select), xlab = "Expected Value", ylab = "Residual", main = "")
qqline(resid(lm_select))
title("(a) QQ Plot for Y (trans_losdays2)")
```

###All diagnostics pltos
```{r}
par(mfrox=c(2,2))
plot(lm_select)
```

###Find the outliers
```{r}
stu_res<-rstandard(lm_select)
outliers_y<-stu_res[abs(stu_res)>2.5]
```


###Remove extreme observations 
```{r}
hospital_no_outliers<-data_update[c(-8,-41,-133,-138,-189,-213,-233,-334,-508,-800,-818,-824,-832,-849,-910,-945,-1065,-1068,-1116,-1148,-1168,-1171,-1197,-1209,-1369,-1429,-1473,-1674,-1712,-1731,-1829,-1839,-1878,-12019,-2065,-2113,-2152,-2201,-2232,-2257,-2288,-2429,-2481),]
mult.fit_no_outliers<- lm(trans_losdays2 ~ is30dayreadmit + cindex_new + evisit_new + ageyear + marital_new + insurance_new + respirationrate +  bpdiastolic + bpsystolic + temperature + heartrate, data=hospital_no_outliers) 

summary(mult.fit_no_outliers)
#After removing, the r-squared value increases by approximately 0.2.
```
#####The reason We remove these outliers is not because these are outliers but we have to meet the normality assumptions.

###Checking for multicollinearity: VIF
```{r}
vif(mult.fit_no_outliers)
#All the values are far below 5, so it indicates there's no collinearity here.
```

###Bootstrap: the result of the estimates seems quite acceptable, low bias
```{r}
set.seed(1)

# Our usual regression, no bootstrap yet
boot.fn<-function(data, index){
	return(coef(lm(trans_losdays2 ~ is30dayreadmit + cindex_new + evisit_new + ageyear + marital_new + insurance_new + respirationrate +  bpdiastolic + bpsystolic + temperature + heartrate,data=data,subset=index)))
}
boot.fn(hospital_no_outliers,1:2546)



# Use function boot() to repeat the sampling 10000 times.
# Repeat 10000 times to get the estimates, SEs ad bias

boot(hospital_no_outliers, boot.fn, 10000)

```

###Cross validation: LOOCV
```{r}

# Use glm() instead of lm() because of cv.glm() function
glm.fit<-glm(trans_losdays2 ~ is30dayreadmit + cindex_new + evisit_new + ageyear + marital_new + insurance_new + respirationrate +  bpdiastolic + bpsystolic + temperature + heartrate, data=hospital_no_outliers)

cv.err<-cv.glm(hospital_no_outliers, glm.fit)

cv.err$delta
anova(mult.fit_no_outliers)
# The two delta values are similar, and comparing to the MSE, they are also similar, which means our model has a good predictive ability. However, the r-squared is quite low. So it's quite contradictory..?
summary(glm.fit) #added a summary of the final model
```